---
title: "Missing data update"
author: "Matt T. Trentman"
date: "5/19/2021"
output:
  pdf_document: default
  html_document: default
---
```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
library(rstan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
library(dplyr)
library(shinystan)
library(faux)
library(bayesplot)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(lubridate)
library(data.table)
library(Amelia)
library(tidyr)
library(MASS)
library(plyr)
library(ggplot2)
library(summarytools)
library(knitr)
library(abind)

options(scipen=999)

#setwd("C:/Users/mtrentman/IDrive-Sync/Postdoc/Estimating missing data/daily_predictions")
setwd("C:/Users/matt/IDrive-Sync/Postdoc/Estimating missing data/daily_predictions")
#setwd("~/GitHub/missing-data")
sp<-read.csv('daily.predictions.filled.csv',header = TRUE) ##Missing dates filled in
sp$date.f<-as.Date(sp$date.f,format="%Y-%m-%d")
sp$site_name<-as.character(sp$site_name)
#sp<-read.table(file = 'daily_predictions.tsv',header = TRUE) ##Raw data
sd<-read.table(file = 'site_data.tsv',header = TRUE)##Site data
sd$site_name<-as.character(sd$site_name)

##Function to switch font color depending on the output (HTML or PDF)
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, 
      x)
  } else x
}

```
\
\

### Model
Simulated data using an AR(1) missing data process-model with an intercept, light covariate, and process error.
$$
GPP=\beta_0+\phi\times GPP_{t-1}+\beta_1\times X_{light}+\epsilon_{sdp}
$$
\
\

### Bias 
Bias is calculated as the known parameter used to simulate the data minus the mean parameter estimate of the posterior.

$$
Bias= Parameter_{known}-Mean \; Parameter_{estimated}
$$
\
\

### Relative Width of Credible Interval (RWCI)
The relative width of the credible interval can show how the change in the credible interval with increasing missing data relative to the complete data. 

$$
RWCI=  log\left(\frac{Upper\: CI_{estimated}- Lower\:CI_{estimated}} {Upper\: CI_{complete}- Lower\:CI_{complete}}\right)
$$

\
where:
$$
\begin{aligned}
Upper \: CI&= 97.5\%\: quantile\: of \:posterior \:distribution\\
Lower\: CI&= 2.5\% \: quantile \:of\: posterior \:distribution\\
Estimated &= quantiles\: estimated\: from\: dataset\: with \:missing \:data\\
Complete &= quantiles\: estimated\: from\: complete \:dataset\\
\end{aligned}
$$
\

---

\

## Higher sdp (0.1) and phi (0.8)

### Bayes parameters
``` {r , echo=FALSE, fig.cap=" ", fig.show="hold", out.width="50%"}
##Load model output
fit_sim_summary_bayes <- readRDS("C:/Users/matt/IDrive-Sync/Postdoc/Estimating missing data/daily_predictions/summary_sim_day_bayes_sdp_1_phi_8_b0_1_b1_1.RDS")
known.data.sim<- readRDS("C:/Users/matt/IDrive-Sync/Postdoc/Estimating missing data/daily_predictions/known_sim_day_bayes_sdp_1_phi_8_b0_1_b1_1.RDS")

##Fix known values 
fit_sim_summary_bayes$known.param<-rep(c(0.1, 0.8, 0.1, 0.1), times=8)

##Calculate denominator of rwci for easier calculation
fit_sim_summary_bayes$rwci.den<-rep(fit_sim_summary_bayes$high[1:4]-fit_sim_summary_bayes$min[1:4], times=8)


##Calculate bias (difference from known) and rwci (relative width of the credible interval)

diff<-ddply(fit_sim_summary_bayes, c("prop.missing", "param"),summarize,
             bias=mean-known.param, 
             rwci= (high-min)/rwci.den
             )

fit_sim_summary_bayes$bias<-diff$bias
fit_sim_summary_bayes$rwci<-diff$rwci


ggplot(data=fit_sim_summary_bayes, aes(x=bias, y=prop.missing ))+
  geom_point(aes( color=param, group=param),size=3,position=position_dodge(0.5))+
  theme_classic()+
  theme(legend.position="top")+
  ylab("Percent of Missing Data")+
  xlab("Bias")+
  scale_color_manual(values=c("blue", "green", "darkgray", "black"))+
  scale_x_continuous(limits=c(-0.6,0.35))+
  theme(axis.title.x=element_text(size=18,colour = "black"))+
  theme(axis.title.y=element_text(size=18,colour = "black"))+
  theme(axis.text.y=element_text(size=18,colour = "black"))+
  theme(axis.text.x=element_text(size=18,colour = "black"))+
  geom_vline(xintercept = 0,color=c("black"))

ggplot(data=fit_sim_summary_bayes, aes(x=log(rwci), y=prop.missing ))+
  geom_point(aes( color=param, group=param),size=3,position=position_dodge(0.5))+
  theme_classic()+
  theme(legend.position="top")+
  ylab("Percent of Missing Data")+
  xlab("RWCI")+
  scale_color_manual(values=c("blue", "green", "darkgray", "black"))+
  scale_x_continuous(limits=c(0,1.5))+
  theme(axis.title.x=element_text(size=18,colour = "black"))+
  theme(axis.title.y=element_text(size=18,colour = "black"))+
  theme(axis.text.y=element_text(size=18,colour = "black"))+
  theme(axis.text.x=element_text(size=18,colour = "black"))+
  geom_vline(xintercept = 0,color=c("black"))

```
\

### Amelia parameters
``` {r , echo=FALSE, fig.cap="", fig.show="hold", out.width="50%"}
fit_sim_summary_amelia <- readRDS("C:/Users/matt/IDrive-Sync/Postdoc/Estimating missing data/daily_predictions/summary_sim_day_amelia_sdp_1_phi_8_b0_1_b1_1.RDS")

known.data.sim<- readRDS("C:/Users/matt/IDrive-Sync/Postdoc/Estimating missing data/daily_predictions/known_sim_day_amelia_sdp_1_phi_8_b0_1_b1_1.RDS")


##Create known parameters with some variation for graphing purposes
known.data.sim$known.data<-c(0.1, 0.8, 0.098, 0.102)

##Average across imputations (for now)
summ<-ddply(fit_sim_summary_amelia, c("param", "prop.missing"), summarize,
       avg    = mean(mean),
       ci.min = mean(min),
       ci.max = mean(high))

##Fix known values 
summ$known.param<-rep(c(0.1, 0.1, 0.8, 0.1), each=8)

##Calculate  rwci 
summ<-summ[order(summ$prop.missing),]
summ$rwci.den<-rep(summ$ci.max[1:4]-summ$ci.min[1:4], times=8)
summ$rwci= (summ$ci.max-summ$ci.min)/summ$rwci.den


##Calculate bias
summ$bias=(summ$avg-summ$known.param)


ggplot(data=summ, aes(x=bias, y=prop.missing ))+
    geom_point(aes( color=param, group=param),size=3,position=position_dodge(0.5))+
    theme_classic()+
    theme(legend.position="top")+
    ylab("Percent of Missing Data")+
    xlab("Bias")+
    scale_color_manual(values=c("blue", "green", "darkgray", "black"))+
    scale_x_continuous(limits=c(-0.6,0.35))+
    theme(axis.title.x=element_text(size=18,colour = "black"))+
    theme(axis.title.y=element_text(size=18,colour = "black"))+
    theme(axis.text.y=element_text(size=18,colour = "black"))+
    theme(axis.text.x=element_text(size=18,colour = "black"))+
    geom_vline(xintercept = 0,color=c("black"))


ggplot(data=summ, aes(x=log(rwci), y=prop.missing ))+
    geom_point(aes( color=param, group=param),size=3,position=position_dodge(0.5))+
    theme_classic()+
    theme(legend.position="top")+
    ylab("Percent of Missing Data")+
    xlab("RWCI")+
    scale_color_manual(values=c("blue", "green", "darkgray", "black"))+
    scale_x_continuous(limits=c(0,1.5))+
    theme(axis.title.x=element_text(size=18,colour = "black"))+
    theme(axis.title.y=element_text(size=18,colour = "black"))+
    theme(axis.text.y=element_text(size=18,colour = "black"))+
    theme(axis.text.x=element_text(size=18,colour = "black"))+
    geom_vline(xintercept = 0,color=c("black"))


```
\

---

\

## Lower sdp (0.01) and phi (0.8)

### Bayes parameters
``` {r , echo=FALSE, fig.cap=" ", fig.show="hold", out.width="50%"}

fit_sim_summary_bayes <- readRDS("C:/Users/matt/IDrive-Sync/Postdoc/Estimating missing data/daily_predictions/summary_sim_day_bayes_sdp_01_phi_8_b0_1_b1_1.RDS")
known.data.sim<- readRDS("C:/Users/matt/IDrive-Sync/Postdoc/Estimating missing data/daily_predictions/known_sim_day_bayes_sdp_01_phi_8_b0_1_b1_1.RDS")

##Fix known values 
fit_sim_summary_bayes$known.param<-rep(c(0.01, 0.8, 0.102, 0.1), times=8)

##Calculate denominator of rwci for easier calculation
fit_sim_summary_bayes$rwci.den<-rep(fit_sim_summary_bayes$high[1:4]-fit_sim_summary_bayes$min[1:4], times=8)


##Calculate bias (difference from known) and rwci (relative width of the credible interval)

diff<-ddply(fit_sim_summary_bayes, c("prop.missing", "param"),summarize,
             bias=mean-known.param, 
             rwci= (high-min)/rwci.den
             )

fit_sim_summary_bayes$bias<-diff$bias
fit_sim_summary_bayes$rwci<-diff$rwci





ggplot(data=fit_sim_summary_bayes, aes(x=bias, y=prop.missing ))+
  geom_point(aes( color=param, group=param),size=3,position=position_dodge(0.5))+
  theme_classic()+
  theme(legend.position="top")+
  ylab("Percent of Missing Data")+
  xlab("Bias")+
  scale_color_manual(values=c("blue", "green", "darkgray", "black"))+
  scale_x_continuous(limits=c(-0.5,0.35))+
  theme(axis.title.x=element_text(size=18,colour = "black"))+
  theme(axis.title.y=element_text(size=18,colour = "black"))+
  theme(axis.text.y=element_text(size=18,colour = "black"))+
  theme(axis.text.x=element_text(size=18,colour = "black"))+
  geom_vline(xintercept = 0,color=c("black"))

ggplot(data=fit_sim_summary_bayes, aes(x=log(rwci), y=prop.missing ))+
  geom_point(aes( color=param, group=param),size=3,position=position_dodge(0.5))+
  theme_classic()+
  theme(legend.position="top")+
  ylab("Percent of Missing Data")+
  xlab("RWCI")+
  scale_color_manual(values=c("blue", "green", "darkgray", "black"))+
  scale_x_continuous(limits=c(-.05,1.5))+
  theme(axis.title.x=element_text(size=18,colour = "black"))+
  theme(axis.title.y=element_text(size=18,colour = "black"))+
  theme(axis.text.y=element_text(size=18,colour = "black"))+
  theme(axis.text.x=element_text(size=18,colour = "black"))+
  geom_vline(xintercept = 0,color=c("black"))


```
\

### Amelia parameters
``` {r , echo=FALSE, fig.cap="", fig.show="hold", out.width="50%"}
fit_sim_summary_amelia <- readRDS("C:/Users/matt/IDrive-Sync/Postdoc/Estimating missing data/daily_predictions/summary_sim_day_amelia_sdp_01_phi_8_b0_1_b1_1.RDS")
known.data.sim<- readRDS("C:/Users/matt/IDrive-Sync/Postdoc/Estimating missing data/daily_predictions/known_sim_day_amelia_sdp_01_phi_8_b0_1_b1_1.RDS")

##Create known parameters with some variation for graphing purposes
known.data.sim$known.data<-c(0.01, 0.8, 0.1, 0.102)

##Average across imputations (for now)
summ<-ddply(fit_sim_summary_amelia, c("param", "prop.missing"), summarize,
       avg    = mean(mean),
       ci.min = mean(min),
       ci.max = mean(high))

##Fix known values 
summ$known.param<-rep(c(0.1, 0.1, 0.8, 0.01), each=8)

##Calculate  rwci 
summ<-summ[order(summ$prop.missing),]
summ$rwci.den<-rep(summ$ci.max[1:4]-summ$ci.min[1:4], times=8)
summ$rwci= (summ$ci.max-summ$ci.min)/summ$rwci.den


##Calculate bias
summ$bias=(summ$avg-summ$known.param)



ggplot(data=summ, aes(x=bias, y=prop.missing ))+
    geom_point(aes( color=param, group=param),size=3,position=position_dodge(0.5))+
    theme_classic()+
    theme(legend.position="top")+
    ylab("Percent of Missing Data")+
    xlab("Bias")+
    scale_color_manual(values=c("blue", "green", "darkgray", "black"))+
    scale_x_continuous(limits=c(-0.5,0.35))+
    theme(axis.title.x=element_text(size=18,colour = "black"))+
    theme(axis.title.y=element_text(size=18,colour = "black"))+
    theme(axis.text.y=element_text(size=18,colour = "black"))+
    theme(axis.text.x=element_text(size=18,colour = "black"))+
    geom_vline(xintercept = 0,color=c("black"))


ggplot(data=summ, aes(x=log(rwci), y=prop.missing ))+
    geom_point(aes( color=param, group=param),size=3,position=position_dodge(0.5))+
    theme_classic()+
    theme(legend.position="top")+
    ylab("Percent of Missing Data")+
    xlab("RWCI")+
    scale_color_manual(values=c("blue", "green", "darkgray", "black"))+
    scale_x_continuous(limits=c(0,1.5))+
    theme(axis.title.x=element_text(size=18,colour = "black"))+
    theme(axis.title.y=element_text(size=18,colour = "black"))+
    theme(axis.text.y=element_text(size=18,colour = "black"))+
    theme(axis.text.x=element_text(size=18,colour = "black"))+
    geom_vline(xintercept = 0,color=c("black"))





```

