---
title: "State-space model"
author: "Matt T. Trentman"
date: "11/23/2021"
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
library(rstan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
library(dplyr)
library(shinystan)
library(faux)
library(bayesplot)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(lubridate)
library(data.table)
library(Amelia)
library(tidyr)
library(MASS)
library(plyr)
library(ggplot2)
library(summarytools)
library(knitr)
library(abind)

options(scipen=999)

```
# Goal
Create a simple state-space model that recovers known parameters from simulated GPP 'data'.
My ideal model is an auto-regressive process model with an intercept ($\beta_0$), AR(1) term ($\phi$), light covariate ($\beta_1$), and process error ($\sigma_{sdp}$) that are used to predict the state ($GPP_{state,\ t}$) at time $t$.
The observation model includes the estimated state and observation error $(\sigma_{obs})$
$$
\begin{aligned}
GPP_{state,\ t}=\beta_0+\phi\times GPP_{state,\ t-1}+\beta_1\times X_{light}+\sigma_{sdp}\\
GPP_{obs}=GPP_{state}+\sigma_{obs}
\end{aligned}
$$


# The Problem
When I simulate data using the above equations, I cannot recover the known parameters. There seems to be an issue with $\sigma_{obs}$ and $\sigma_{sdp}$ in particular. 
\newpage

# What I have done so far
## Process-only model
On Bob's suggestion, I backtracked to a simpler model to see where things go wrong. 
I started with a simple process error only model with autocorrelation where:
\
$\beta_0$=0\
$\sigma_{sdp}$= 0.01\
$\phi$=0.8\
$$
GPP_{state,\ t}=\beta_0+\phi\times GPP_{state,\ t-1}+\sigma_{sdp}\\
$$


```{r fig1, echo=FALSE, fig.cap="Process-only model simulated data ", fig.align = "center", out.width='70%'}
#Simulate data

N<-365 #length of data
t<-1:N #time
time<-seq(as.POSIXct("2020/1/1"),as.POSIXct("2020/12/30"), by="day") #for light

##GPP based on time-series model with known parameters
set.seed(550)
z<-NA
sdp <- 0.01
phi <-0.8
z[1]<-0

## Set the seed, so we can reproduce the results
set.seed(550)
## For-loop that simulates the state through time, using i instead of t,
for (t in 2:N){
  z[t] = phi*z[t-1]+rnorm(1, 0, sdp)
}

y_full<-data.frame(z,t)

ggplot(data=y_full, aes(x=time, y=z))+
  geom_point(color="black", size=3)+
  theme_classic()+
  theme(axis.title.x=element_text(size=18,colour = "black"))+
  theme(axis.title.y=element_text(size=18,colour = "black"))+
  theme(axis.text.y=element_text(size=18,colour = "black"))+
  theme(axis.text.x=element_text(size=18,colour = "black"))+
  theme(legend.position="top")+
  ylab("Data")+
  xlab("Time (days)")+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```
\newpage

### STAN Code and output
This code uses bits from Auger-Methe et al. (2021, Ecological Monigraphs) and various blogs on state-space models in Stan.
```{r eval=FALSE}
#STAN model
sink("process.stan")

cat(
 
/*----------------------- Data --------------------------*/
  data {
    int N; // Length of state and observation time series
    vector[N] z; // Observations
    real z0;
    }
  
/*----------------------- Parameters --------------------------*/
  
  parameters {
    real<lower=0> sdp; // Standard deviation of the process equation
    real b0;
    real<lower = 0, upper=1 > phi; // Auto-regressive parameter
      }
  

  /*----------------------- Model --------------------------*/
  model {
    // Prior distributions
    sdp ~ normal(0, 1);
    phi ~ beta(1,1);
    b0 ~ normal(0,5);
    
    // Distribution for the first state
    z[1] ~ normal(z0, sdp);
  
    // Distributions for all other states
    for(t in 2:N){
       z[t] ~ normal(b0+phi*z[t-1], sdp);// process model with error
    }
  }
  
    ,fill=TRUE)
sink()

#Prep model
model<-"process.stan"
model<-stan_model(model)

#Create data object
data <- list(N = length(z),z=z, z0=z[1])

#Run Stan
fit<-rstan::sampling(object=model, data = data,  iter = 3000, chains = 4)
```
\newpage

```{r fig2, echo=FALSE, fig.cap=c("Traceplot of process-only model", "Posterior distribution of process-only model with known values"), fig.align = "center", out.width='60%'}
##Traceplots
fit<-readRDS("C:/Users/matt/Documents/GitHub/missing-data/process.RDS")

traceplot(fit, pars=c("phi", "sdp", "b0"))

##Density plots
plot_sdp <- stan_dens(fit, pars="sdp") + geom_vline(xintercept = sdp)
plot_phi <- stan_dens(fit, pars="phi") + geom_vline(xintercept = phi)
plot_b0 <- stan_dens(fit, pars="b0") + geom_vline(xintercept = 0)
grid.arrange(plot_phi,plot_sdp,plot_b0, nrow=2)
```
The chains look great, and the known parameters match fairly well with the posterior distributions. **I am not sure if this is 'good enough' or if we should try for more accuracy**. For now, lets move on to a state-space model (add observation error).
\newpage

## Simple State-Space model
Add an observation model equation to the process model above.
$$
\begin{aligned}
GPP_{state,\ t}=\beta_0+\phi\times GPP_{state,\ t-1}+\sigma_{sdp}\\
GPP_{obs}=GPP_{state}+\sigma_{obs}
\end{aligned}
$$
Simulate some data-both the 'true/latent state', which only includes process error, as well as the observed data, which is the 'true/latent state' but with observation error. Here are the parameters I used for the model:\
$\beta_0$=0\
$\sigma_{sdp}$= 0.01\
$\sigma_{sdo}$= 0.01\
$\phi$=0.8\

```{r echo=FALSE, fig.cap="Simulated data for state-space model", align='center', out.width='80%'}
#Simulate data

N<-500#length of data
z<-numeric(N+1) 
t<-numeric(N+1) 
t<-0:500
##GPP based on time-series model with known parameters
set.seed(553)
sdp <- 0.01
sdo<-0.01
phi<-.8

## Set the seed, so we can reproduce the results
set.seed(550)
sd.p<-rnorm(N, 0, sdp)
## For-loop that simulates the state through time
for (i in 1:N){
  z[i+1] =z[i]*phi+sd.p[i]
}

#Estimate observed data
set.seed(500)
sd.o <-rnorm(N, 0, sdo)
y <-z[2:(N+1)]+ sd.o

##Plot
plot(1:length(y), y,
     pch=3, cex = 0.8, col="blue", ty="o", lty = 3,
     xlab = "t", ylab = expression(y[t]),
     xlim = c(0,N), #ylim = c(min(y), max(y+max(y)/5)),
     las = 1)
points(0:N, z,
       pch = 19, cex = 0.7, col="red", ty = "o")
legend("top",
       legend = c("Obs.", "True states"),
       pch = c(3, 19),
       col = c("blue", "red"),
       lty = c(3, 1),
       horiz=TRUE, bty="n", cex=0.9)

```
\newpage
### An aside
I have found that there are many ways to simulated the two data sets, and it is easy to simulate the data incorrectly. I was doing this wrong for a long time...

I originally used this set of equations:
```{r eval=FALSE}
set.seed(550)
## For-loop that simulates the latent state through time
for (i in 2:N){
  z[i] =z[i-1]+rnorm(1, 0, sdp)
}

#Estimate observed data
y <-z+rnorm(N, 0, sdo)
```

But these equations create an offset of error between the two simulated vectors. I used Auger-Methes (2021) approach and it fixed the problem. Long-story short, you need to estimate time point 0 for the state to start the simulation. Then apply the observation error to index 2 of the state, which is actually time point 1.
```{r eval=FALSE}
## For-loop that simulates the state through time
for (i in 1:N){
  z[i+1] =z[i]*phi+rnorm(1, 0, sdp)
}

#Estimate observed data
y <-z[2:(N+1)]+ rnorm(N, 0, sdo)
```

\newpage

### STAN Code and output
Now onto the STAN code and output for the state-space model.
```{r eval=FALSE}
#STAN model
sink("ss_1.stan")

cat("
 

/*----------------------- Data --------------------------*/
  data {
    int N; // Length of state and observation time series
    vector[N] y; // Observations
    real z0; //intital state value
    }
  
/*----------------------- Parameters --------------------------*/
  
  parameters {
    vector[N] z; //latent state variable
    real<lower=0> sdp; // Standard deviation of the process equation
    real<lower=0> sdo; // Standard deviation of the observation equation
    real<lower = 0, upper=1 > phi; // Auto-regressive parameter
    real b0; 
      }
  

  /*----------------------- Model --------------------------*/
  model {
    // Prior distributions
    sdo ~ normal(0, 1);
    sdp ~ normal(0, 1);
    phi ~ beta(1,1);
    b0  ~ normal(0,5);
    
    // Distribution for the first state
    z[1] ~ normal(z0,sdp);
  
    // Distributions for all other states
    for(i in 2:N){
       z[i] ~ normal(z[i-1]*phi+b0, sdp);// process model with error
    }
    
    for(i in 1:N){
       y[i] ~ normal(z[i], sdo); // observation model with fixed observation error
    }

 }
"
    ,fill=TRUE)
sink()
closeAllConnections()

#Prep model
model<-"ss_1.stan"
model<-stan_model(model)

#Create data object
data <- list(N = length(z),z=z, z0=z[1])

#Run Stan
fit<-rstan::sampling(object=model, data = data,  iter = 3000, chains = 4)
```

\newpage

```{r echo=FALSE, fig.cap=c("Traceplot of state-space model", "Posterior distribution of state-space model with known values"), fig.align = "center", out.width='60%'}
fit<-readRDS("C:/Users/matt/Documents/GitHub/missing-data/state_space_1.RDS")

##Traceplots
traceplot(fit, pars=c("sdo", "sdp", "phi", "b0"))

##Plot density plots
plot_sdo <- stan_dens(fit, pars="sdo") + geom_vline(xintercept =sdo)
plot_sdp <- stan_dens(fit, pars="sdp") + geom_vline(xintercept = sdp)
plot_phi <- stan_dens(fit, pars="phi") + geom_vline(xintercept = phi)
plot_b0 <- stan_dens(fit, pars="b0") + geom_vline(xintercept =0)
grid.arrange(plot_sdp,plot_sdo,plot_phi, plot_b0,nrow=2)
```
Again, these look pretty good. No errors, chains are stationary and well mixed, and the known parameters match with the posterior distribution (although, again, it's unclear to me if this is *good enough*.

\newpage

## The Rub with the state-space model
Issues form when we **change the value of sdp or sdo**. If they are the same value it works fine (even at different magnitudes; say sdp/sdo=c(0.001,0.01,0.1)). Things fall apart if they are different.
For example, lets set:\
$\beta_0$=0\
$\sigma_{sdp}$= 0.01\
$\sigma_{sdo}$= 0.001\
$\phi$=0.8\

And use the same state-space model as above.

```{r echo=FALSE, fig.cap=" ", fig.show="hold"}
#Simulate data

N<-500#length of data
z<-numeric(N+1) 
t<-numeric(N+1) 
t<-0:500
##GPP based on time-series model with known parameters
set.seed(553)
sdp <- 0.01
sdo<-0.001
phi<-.8

## Set the seed, so we can reproduce the results
set.seed(550)
sd.p<-rnorm(N, 0, sdp)
## For-loop that simulates the state through time
for (i in 1:N){
  z[i+1] =z[i]*phi+sd.p[i]
}

#Estimate observed data
set.seed(500)
sd.o <-rnorm(N, 0, sdo)
y <-z[2:(N+1)]+ sd.o


##Plot
plot(1:length(y), y,
     pch=3, cex = 0.8, col="blue", ty="o", lty = 3,
     xlab = "t", ylab = expression(y[t]),
     xlim = c(0,N), #ylim = c(min(y), max(y+max(y)/5)),
     las = 1)
points(0:N, z,
       pch = 19, cex = 0.7, col="red", ty = "o")
legend("top",
       legend = c("Obs.", "True states"),
       pch = c(3, 19),
       col = c("blue", "red"),
       lty = c(3, 1),
       horiz=TRUE, bty="n", cex=0.9)
```
\newpage

```{r echo=FALSE, fig.cap=c("Traceplot of state-space model", "Posterior distribution of state-space model with known values"), fig.align = "center", out.width='60%'}
fit<-readRDS("C:/Users/matt/Documents/GitHub/missing-data/state_space_2.RDS")

##Traceplots
traceplot(fit, pars=c("sdo", "sdp", "phi", "b0"))

##Plot density plots
plot_sdo <- stan_dens(fit, pars="sdo") + geom_vline(xintercept =0.001)
plot_sdp <- stan_dens(fit, pars="sdp") + geom_vline(xintercept = 0.01)
plot_phi <- stan_dens(fit, pars="phi") + geom_vline(xintercept = phi)
plot_b0 <- stan_dens(fit, pars="b0") + geom_vline(xintercept =0)
grid.arrange(plot_sdp,plot_sdo,plot_phi, plot_b0,nrow=2)
```
Woah. Chaos. There are some issues with estimating sdo and sdp.

\newpage

Even worse if we flip the values for sdp/sdo!
\
$\beta_0$=0\
$\sigma_{sdp}$= 0.001\
$\sigma_{sdo}$= 0.01\
$\phi$=0.8\

```{r echo=FALSE, fig.cap=c("Traceplot of state-space model", "Posterior distribution of state-space model with known values"), fig.align = "center", out.width='60%'}
fit<-readRDS("C:/Users/matt/Documents/GitHub/missing-data/state_space_3.RDS")

##Traceplots
traceplot(fit, pars=c("sdo", "sdp", "phi", "b0"))

##Plot density plots
plot_sdo <- stan_dens(fit, pars="sdo") + geom_vline(xintercept =0.01)
plot_sdp <- stan_dens(fit, pars="sdp") + geom_vline(xintercept = 0.001)
plot_phi <- stan_dens(fit, pars="phi") + geom_vline(xintercept = 0.8)
plot_b0 <- stan_dens(fit, pars="b0") + geom_vline(xintercept =0)
grid.arrange(plot_sdp,plot_sdo,plot_phi, plot_b0,nrow=2)
```
\newpage

## My (unsuccessful) attempts to fix the problem
* Hagiwara 2021- Time Series Analysis for the State-Space Model with R/Stan) does a few things different than our model. I tried a few and they don't help.   + modeling sdo and sdo as covariance matrices
  + modeling z0 as a parameter to inform z1
* Increasing sample size from 500-->5000
* Strong prior on sdo (sdo ~ normal(0, 0.001))

\newpage

## Somewhat successful attempts
* Fixing $\sigma_{sdo}$= to the true value
 + $\beta_0$=0\
$\sigma_{sdp}$= 0.5\
$\sigma_{sdo}$= 0.01\
$\phi$=0.8\

```{r echo=FALSE, fig.cap=c("Traceplot of state-space model", "Posterior distribution of state-space model with known values"), fig.align = "center", out.width='60%'}
fit<-readRDS("C:/Users/matt/Documents/GitHub/missing-data/ss_AM_fixedsdo.RDS")

##Traceplots
traceplot(fit, pars=c( "sdp", "phi", "b0"))

##Plot density plots
plot_sdp <- stan_dens(fit, pars="sdp") + geom_vline(xintercept =0.5)
plot_phi <- stan_dens(fit, pars="phi") + geom_vline(xintercept =0.8)
plot_b0 <- stan_dens(fit, pars="b0") + geom_vline(xintercept =0)
grid.arrange(plot_sdp,plot_phi, plot_b0,nrow=2)
```


